---
title: "204problemset2"
author: "Claire Madden"
date: "4/25/2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#set up 
library(tidyverse)
library(rootSolve)
library(reconPlots)


#read in data
gasdata <- read_csv("Gas_Data.csv")

#change column headers to be more simple
names(gasdata) <- c("price", "q_low", "q_high")


```


1. What is the aggregate daily demand curve for gasoline in France? What is the supply curve for gasoline? What is the “benefit” to consumers under the status quo? What is the “benefit” to producers under the status quo? What is the environmental cost under the status quo?

```{r}
#need to find: high demand curve, low demand curve, aggregate demand curve, supply curve (mpc), environmental cost curve (mec) 

#high linear model demand curve:

d_high <- lm(price ~ q_high, data = gasdata)
summary(d_high)
#p=a(q)+b
#p= -0.0001297(q) + 23.39

#saving output numbers for use later
a_high <- d_high$coefficients[2]
b_high <- d_high$coefficients[1]

#high demand curve function p=a*q+b
demand_high <- function(x){a_high*x + b_high}

#variables for high demand curve q=p/a - b/a
a_high1 <- (1/a_high)
b_high1 <- (-b_high/a_high)


#low demand curve:

d_low <- lm(price ~ q_low, data = gasdata)
summary(d_low)
#p=a(q)+b
#p= -0.0001355(q) + 21.99

#saving output numbers for use later
a_low <- d_low$coefficients[2]
b_low <- d_low$coefficients[1]

#low demand curve function p=a*q+b
demand_low <- function(x){a_low*x + b_low}

#variables for low demand curve q=p/a - b/a
a_low1 <- (1/a_low)
b_low1 <- (-b_low/a_low)


#combine high and low variables to get aggregate 
a_agg <- a_high1 + a_low1
b_agg <- b_high1 + b_low1

#rearrage to get p=
a_agg1 <- (1/a_agg)
b_agg1 <- (-b_agg/a_agg)

agg_demand <- function(x){a_agg1*x + b_agg1}

#need to know what q equals when p is 21.99085 for aggregate demand curve


#supply curve (mpc)
supply <- 5/((5-b_agg1)/(a_agg1))

#supply function
supply_curve <- function(x){supply*x}

#environmental cost
mec <- function(x){0*x + 2}

#social cost
msc <- function(x){supply*x + 2}


#graph both demand functions on the same axes
allcurves <- ggplot(data.frame(x = c(7000, 500000)), aes(x = x)) +
  stat_function(fun = demand_low, color = "cadetblue4") +
  stat_function(fun = demand_high, color = "orchid4")+
  stat_function(fun = agg_demand, color = "tomato3")+
  stat_function(fun = mec, color = "turquoise")+
  stat_function(fun = msc, color = "gold")+
  stat_function(fun = supply_curve, color = "olivedrab4")+
  scale_y_continuous(limits = c(0,30), expand = c(0,0))+
  scale_x_continuous(limits = c(0,400000), expand = c(0,0))


allcurves

```

```{r}

x_range <- 0:500000

#intersection of aggregate demand and high demand curves
kink <- curve_intersect(agg_demand, demand_high, empirical = FALSE, domain = c(min(x_range), max(x_range))) 

kink

kink_q <- kink$x

kink_p <- kink$y

#Sam Update: intersection of agg demand and supply - carry over price to find quantity for Qh and Ql 


#intersection of supply and agg demand

int_supp_aggd <- curve_intersect(supply_curve, agg_demand, empirical = FALSE, domain = c(min(x_range), max(x_range)))

int_supp_aggd

int_supp_aggd_q <- int_supp_aggd$x

int_supp_aggd_p <- int_supp_aggd$y

# low quantity at equlibrium price

low_q <- ((5-b_low)/a_low)

low_q

# high quantity at equlibrium price

high_q <- ((5-b_high)/a_high)

high_q


#intersection of msc and agg demand

int_msc_aggd <- curve_intersect(msc, agg_demand, empirical = FALSE, domain = c(min(x_range), max(x_range)))

int_msc_aggd
                                
int_msc_aggd_q <- int_msc_aggd$x
#243680.6
int_msc_aggd_p <- int_msc_aggd$y
#6.5596



```


```{r}
#low demand benefit under status quo

base_cs_low <- b_low - int_supp_aggd_p
height_cs_low <- low_q

cs_low <- (1/2)*base_cs_low*height_cs_low

cs_low



#high demand benefit under status quo

base_cs_high <- b_high - int_supp_aggd_p
height_cs_high <- high_q

cs_high <- (1/2)*base_cs_high*height_cs_high

cs_high



#aggregate benefit under status quo

cs_agg <- cs_high + cs_low

cs_agg

# producer surplus

base_ps <- int_supp_aggd_q
height_ps <- int_supp_aggd_p

ps_agg <- (1/2)*base_ps*height_ps

ps_agg


```

```{r}
#environmental cost under status quo

base_tec <- int_supp_aggd_q
height_tec <- 2

tec <- base_tec*height_tec

tec

```

2. How is the current consumer benefit divided between “High” and “Low” income consumers?
```{r}

prop_low <- cs_low/cs_agg
prop_low

prop_high <- cs_high/cs_agg
prop_high

```

3. A gas tax of $0.50/gal. is proposed. What would be the effects of this tax on:
a. The amount of gasoline produced and sold in France.
b. The price of gasoline.
c. Welfare of “High” income consumers.
d. Welfare of “Low” income consumers.
e. Welfare of gas producers.
f. Total environmental damage.
g. Total revenue generated by the tax.


```{r}

supply_tax <- function(x){supply*x + 0.5}

#a. amount of gasoline produced and sold in france
q_tax <- curve_intersect(supply_tax, agg_demand, empirical = FALSE, domain = c(min(x_range), max(x_range)))

q_tax

tax0.5_q <- q_tax$x 

#b. price of gasoline with tax
tax0.5_p <- q_tax$y

tax0.5_p

# low quantity with tax

low_q_tax <- ((5.3899-b_low)/a_low)

# high quantity with tax

high_q_tax <- ((5.3899-b_high)/a_high)

high_q_tax


#c. high cs

base_high_tax0.5 <- high_q_tax
height_high_tax0.5 <- b_high - tax0.5_p

cs_tax_high <- (1/2)*base_high_tax0.5*height_high_tax0.5
cs_tax_high

#d. low cs

base_low_tax0.5 <- low_q_tax
height_low_tax0.5 <- b_low - tax0.5_p

cs_tax_low <- (1/2)*base_low_tax0.5*height_low_tax0.5
cs_tax_low

#e. producer surplus

base_ps_tax0.5 <- tax0.5_q
height_ps_tax0.5 <- tax0.5_p - 0.5

ps_tax0.5 <- (1/2)*base_ps_tax0.5*height_ps_tax0.5
ps_tax0.5

#f. total environmental damage

tec_tax0.5 <- tax0.5_q*2
tec_tax0.5

#g. tax revenue

tax_rev <- tax0.5_q*0.5
tax_rev


```

4. Now, assume that all revenue from a tax will be used for infrastructure repairs, and that the benefit of this is proportional to the amount you drive (i.e. to the gas you consume). tax revenue is considered benefit for each group - +quantity of gas consumed * tax 

Also assume that “Low” income consumers bear all environmental costs. For a range of gas taxes (ranging from $0 - $5.00/gal), calculate the effects of the tax on:
a. Overall welfare of “High” income consumers
b. Overall welfare of “Low” income consumers (CS - tot env damage + tax revenue that they benifit from (proportion that is generated by quanity where supply intercects low demand curve)

c. Gas producers

```{r}

#tax of $2

supply_2tax <- function(x){supply*x + 2}

int_agg_2tax<- curve_intersect(supply_2tax, agg_demand, empirical = FALSE, domain = c(min(x_range), max(x_range)))

int_agg_2tax

int_agg_2tax_p <- int_agg_2tax$y

int_agg_2tax_q <- int_agg_2tax$x

#wellfare high income
# high demand quantity at $2 tax price
high_2tax_q <- ((6.5596 - b_high)/a_high)

high_2tax_q


base_high_2tax <- b_high - int_agg_2tax_p
height_high_2tax <- high_2tax_q

#high cs
cs_high_2tax <- ((1/2)*base_high_2tax*height_high_2tax)+(2*high_2tax_q)

cs_high_2tax


#wellfare low income 

#low demand quantity at $2 tax price
low_2tax_q <- ((6.5596 - b_low)/a_low)

low_2tax_q

base_low_2tax <- b_low - int_agg_2tax_p
height_low_2tax <- low_2tax_q


#environmental cost 
tec_2tax <- int_agg_2tax_q*2

tec_2tax

# low cs 
cs_low_2tax <- ((1/2)*base_low_2tax*height_low_2tax) + (2*low_2tax_q) - tec_2tax

cs_low_2tax

```



5. Finally, assume that electric cars will gain popularity and that in the future this
will lower the demand curves of all income groups by half (vertically). Under these new demand curves, what are the effects on:
a. Gas consumption by “High” income consumers
b. Gas consumption by “Low” income consumers
c. Gas price
d. Environmental damage from gasoline

```{r}

# considering chance WITH $0.5 tax 

#lower demand curves by half vertically

half_demand_high <- function(x){a_high*x + 0.5*b_high}

half_demand_low <- function(x){a_low*x + 0.5*b_low}

half_agg_demand <- function(x){a_agg1*x + 0.5*b_agg1}

#intersect new agg demand and supply with tax

int_tax_half_agg_demand <- curve_intersect(supply_tax, half_agg_demand, empirical = FALSE, domain = c(min(x_range), max(x_range)))

int_tax_half_agg_demand

tax_half_agg_demand_p <- int_tax_half_agg_demand$y

tax_half_agg_demand_p

tax_half_agg_demand_q <- int_tax_half_agg_demand$x

#a. gas consumption by high demand

half_high_demand <- ((2.8899-(0.5*b_high))/a_high)

half_high_demand


#b. gas consumption by low demand

half_low_demand <- ((2.8899-(0.5*b_low))/a_low)
half_low_demand

#c. gas price at new equilibrium - where aggregate demand (reduced by half) intersects supply with tax

tax_half_agg_demand_p

#d. environmental damage from gasoline

#quantity at new equilirium
int_tax_half_agg_demand_q <- int_tax_half_agg_demand$x

# environmental cost at new equilibrium

tec_half_demand <- tax_half_agg_demand_q*2

tec_half_demand

```



6. Using these numbers, which has a bigger impact on environmental quality, a $2.00/gal. gas tax or the invention of electric cars? Provide your evidence.

```{r}

#intersect of supply with $2 tax and agg demand

int_supply_2tax_agg <- curve_intersect(supply_2tax, agg_demand, empirical = FALSE, domain = c(min(x_range), max(x_range)))

int_supply_2tax_agg_q <- int_supply_2tax_agg$x

#environmental cost with $2 tax

tec_2tax <- int_supply_2tax_agg_q*2

tec_2tax
#487361.2

#environmental cost with electric cars

tec_half_demand
#255448.9

```




